---
title: "Decision Tree Classification"
author: "Gayathri Sripathy"
date: "Jan. 23, 2018"
output: 
  html_document:
    number_sections: yes
    theme: readable
    highlight: tango
    toc: yes
    fig_width: 15
    fig_height: 10
---

# Set up, Data import and inspections
## Loading Packages & import data
```{r - Set up, data import and inspections, warning=FALSE, message=FALSE}

# Load packages after they have been installed
# Package loading. Install the following packages before running this chunk or knitting this program.

library(C50)
library(scatterplot3d)
library(caret)
library(rminer)
library(rmarkdown)
library(psych)
library(scatterplot3d)

# Import CD_additional_balanced.csv file
#setwd("C:/Users/gayat/Desktop/Data Mining")
bank <- read.csv(file = "CD_additional_balanced.csv", stringsAsFactors = FALSE)

#Examine the overall data frame
str(bank)
summary(bank)
```

## Changing Character variables to Factor variables
```{r}
# Changing all the nominal/categorical variables into factors. 
bank$job <- factor(bank$job)
bank$marital <- factor(bank$marital)
bank$education <- factor(bank$education)
bank$default <- factor(bank$default)
bank$housing <- factor(bank$housing)
bank$loan <- factor(bank$loan)
bank$contact <- factor(bank$contact)
bank$month <- factor(bank$month)
bank$day_of_week <- factor(bank$day_of_week)
bank$poutcome <- factor(bank$poutcome)
bank$y <- factor(bank$y)

# Re-examine the over data frame 
str(bank)
summary(bank)
```

# Exploring numeric variables
## Histograms - Age, Duration, Campaign and Pdays
```{r}
histogram(bank$age, main="Histogram of Age in the Bank data set",
        xlab="Age")
histogram(bank$duration, main="Histogram of Duration in the Bank data set",
        xlab="Duration")
histogram(bank$campaign, main="Histogram of Campaign in the Bank data set",
        xlab="Campaign")
histogram(bank$pdays, main="Histogram of pdays in the Bank data set",
        xlab="Pdays")
```

## Boxplots - Age, Duration, Campaign and Pdays
```{r}
boxplot(bank$age, main="Boxplot of Age in the Bank data set",
        xlab="Age")
boxplot(bank$duration, main="Boxplot of Duration in the Bank data set",
        xlab="Duration")
boxplot(bank$campaign, main="Boxplot of Campaign in the Bank data set",
        xlab="Campaign")
boxplot(bank$pdays, main="Boxplot of pdays in the Bank data set",
        ylab="Pdays")
```

## Deciles - Age, Duration, Campaign and Pdays
```{r}
quantile(bank$age, seq(from = 0, to = 1, by = 0.10))
quantile(bank$duration, seq(from = 0, to = 1, by = 0.10))
quantile(bank$campaign, seq(from = 0, to = 1, by = 0.10))
quantile(bank$pdays, seq(from = 0, to = 1, by = 0.10))
```

# Exploring factor variables
## Count Value and Percentage value of instances
```{r}
#job, marital, default, education, housing, loan, contact, month, day_of_week, poutcome, y
#the count value and percentage value of instances belonging to that level.
nlevels(bank$job)
Job.table <- table(bank$job)
Job.table
str(Job.table)

nlevels(bank$marital)
Marital.table <- table(bank$marital)
Marital.table
str(Marital.table)

nlevels(bank$default)
Default.table <- table(bank$default)
Default.table
str(Default.table)

nlevels(bank$education)
Education.table <- table(bank$education)
Education.table
str(Education.table)

nlevels(bank$housing)
Housing.table <- table(bank$housing)
Housing.table
str(Housing.table)

nlevels(bank$loan)
Loan.table <- table(bank$loan)
Loan.table
str(Loan.table)

nlevels(bank$contact)
Contact.table <- table(bank$contact)
Contact.table
str(Contact.table)

nlevels(bank$month)
Month.table <- table(bank$month)
Month.table
str(Month.table)

nlevels(bank$day_of_week)
Day.table <- table(bank$day_of_week)
Day.table
str(Day.table)

nlevels(bank$poutcome)
Poutcome.table <- table(bank$poutcome)
Poutcome.table
str(Poutcome.table)

nlevels(bank$y)
y.table <- table(bank$y)
y.table
str(y.table)
```

## Barplots
```{r}
#job, marital, default, education, housing, loan, contact, month, day_of_week, poutcome, y
barplot(Job.table, main = "Barplot of Job in the bank data set", xlab = "Job")
barplot(Education.table, main = "Barplot of Education in the bank data set", xlab = "Education")
barplot(Marital.table, main = "Barplot of Marital in the bank data set", xlab = "Marital")
barplot(Default.table, main = "Barplot of Default in the bank data set", xlab = "Default")
barplot(Housing.table, main = "Barplot of Housing in the bank data set", xlab = "Housing")
barplot(Loan.table, main = "Barplot of Loan in the bank data set", xlab = "Loan")
barplot(Contact.table, main = "Barplot of Contact in the bank data set", xlab = "Contact")
barplot(Month.table, main = "Barplot of Month in the bank data set", xlab = "Month")
barplot(Day.table, main = "Barplot of Day in the bank data set", xlab = "Day")
barplot(Poutcome.table, main = "Barplot of Poutcome in the bank data set", xlab = "Poutcome")
barplot(y.table, main = "Barplot of Y in the bank data set", xlab = "Y")
```

# Exploring relationships amongst multiple variables
## cor and pair.panels for numeric variables
```{r}
#age, duration, campaign, pday, euribor3m, emp.var.rate, and nr.employed
cor(bank[c("age", "duration", "campaign", "pdays", "euribor3m", "emp.var.rate","nr.employed")])
pairs.panels(bank[c("age", "duration", "campaign", "pdays", "euribor3m", "emp.var.rate","nr.employed")])
#duration, emp.var.rate, cons.price.idx, and cons.conf.idx
```

## Boxplots of numeric variables
```{r}
#duration, emp.var.rate, cons.price.idx, and cons.conf.idx
boxplot(duration~y, data = bank)
boxplot(emp.var.rate~y, data = bank)
boxplot(cons.price.idx~y, data = bank)
boxplot(cons.conf.idx~y, data = bank)
```

## Aggregate
```{r}
dur_agg <- aggregate(duration~y, data = bank, FUN = summary)
print(dur_agg)
emp_agg <- aggregate(emp.var.rate~y, summary, data = bank)
print(emp_agg)
price_agg <- aggregate(cons.price.idx~y, summary, data = bank)
print(price_agg)
conf_agg <- aggregate(cons.conf.idx~y, summary, data = bank)
print(conf_agg)
```

## 3d Scatterplot of age, campaign and duration
```{r}
scatterplot3d(bank$age,bank$campaign,bank$duration, pch = as.numeric(bank$y), main = "3D scatter plot of age, campaign and duration variables of bank data")
legend('topright', legend = levels(bank$y),  cex = 0.8, pch = 1:2)
```

## 3d Scatterplot of nr.employed, euribor3m and duration
```{r}
scatterplot3d(bank$nr.employed,bank$euribor3m,bank$duration, pch = as.numeric(bank$y), main = "3D scatter plot of nr.employed, euribor3m and duration variables of bank data")
legend('topright', legend = levels(bank$y),  cex = 0.8, pch = 1:2)
```

# Data preparation
## Data Partition
```{r}
set.seed(100)
inTrain <- createDataPartition(bank$y, p=0.7, list=FALSE)
str(inTrain)
bankTrain <- bank[inTrain,]
bankTest <- bank[-inTrain,]
```

## Distributions in Train and Test data
```{r}
table(bankTrain$y)
table(bankTest$y)
prop.table(table(bankTrain$y))
prop.table(table(bankTest$y))
```

# Train and Test Decision Tree 1 to classify y
## Train a C5.0 model - Decision Tree 1
```{r}
bank_m1_c50 <- C5.0(y~., bankTrain)
bank_m1_c50
#plot(bank_m1_c50)
summary(bank_m1_c50)
```

## predict() and mmetric() functions
```{r}
predicted_bank_test1 <- predict(bank_m1_c50, bankTest)
mmetric(bankTest$y, predicted_bank_test1, metric="CONF")
mmetric(bankTest$y, predicted_bank_test1, metric=c("ACC","TPR","PRECISION","F1"))

predicted_bank_train1 <- predict(bank_m1_c50, bankTrain)
mmetric(bankTrain$y, predicted_bank_train1, metric="CONF")
mmetric(bankTrain$y, predicted_bank_train1, metric=c("ACC","TPR","PRECISION","F1"))
```

# Train and Test Decision Tree 2 to classify y
## Decision Tree 2
```{r}
bank_m2_c50 <- C5.0(bankTrain[c(-21)], bankTrain$y, control = C5.0Control(CF = 0.009))
bank_m2_c50
plot(bank_m2_c50)
summary(bank_m2_c50)
```

## predict() and mmetric() functions
```{r}
predicted_bank_test2 <- predict(bank_m2_c50, bankTest)
mmetric(bankTest$y, predicted_bank_test2, metric="CONF")
mmetric(bankTest$y, predicted_bank_test2, metric=c("ACC","TPR","PRECISION","F1"))

predicted_bank_train2 <- predict(bank_m2_c50, bankTrain)
mmetric(bankTrain$y, predicted_bank_train2, metric="CONF")
mmetric(bankTrain$y, predicted_bank_train2, metric=c("ACC","TPR","PRECISION","F1"))
```

# Train and Test Decision Tree 3 to predict y
## Decision Tree 3
```{r}
#Removing the variable duration
bank_m3_c50 <- C5.0(bankTrain[c(-11,-21)], bankTrain$y)
bank_m3_c50
plot(bank_m3_c50)
summary(bank_m3_c50)
```

## predict() and mmetric() functions
```{r}
predicted_bank_test3 <- predict(bank_m3_c50, bankTest)
mmetric(bankTest$y, predicted_bank_test3, metric="CONF")
mmetric(bankTest$y, predicted_bank_test3, metric=c("ACC","TPR","PRECISION","F1"))

predicted_bank_train3 <- predict(bank_m3_c50, bankTrain)
mmetric(bankTrain$y, predicted_bank_train3, metric="CONF")
mmetric(bankTrain$y, predicted_bank_train3, metric=c("ACC","TPR","PRECISION","F1"))
```

# Training and Testing Decision Tree 4 to classify y
## Decision Tree 4
```{r}
bank_m4_c50 <- C5.0(bankTrain[c(-11,-21)], bankTrain$y, control = C5.0Control(CF = 0.1))
bank_m4_c50
plot(bank_m4_c50)
summary(bank_m4_c50)
```

## predict() and mmetric() functions
```{r}
predicted_bank_test4 <- predict(bank_m4_c50, bankTest)
mmetric(bankTest$y, predicted_bank_test4, metric="CONF")
mmetric(bankTest$y, predicted_bank_test4, metric=c("ACC","TPR","PRECISION","F1"))

predicted_bank_train4 <- predict(bank_m4_c50, bankTrain)
mmetric(bankTrain$y, predicted_bank_train4, metric="CONF")
mmetric(bankTrain$y, predicted_bank_train4, metric=c("ACC","TPR","PRECISION","F1"))
```
